\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-12-11 smurf_paper.tex
should have put actual files instead of links in 32b1a18cd

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers in each of two simultaneous
  imaging bands centred over 450 and 850\,\micron. The camera is
  operated by scanning across the sky and recording data at a rate of
  200\,Hz. As the largest of a new generation of multiplexed kilopixel
  bolometer cameras operating in the (sub)millimetre, \scuba\ data
  analysis represents a significant challenge.  We describe the
  production of maps using the Sub-Millimetre User Reduction Facility
  (SMURF) in which we have adopted a fast, iterative approach to
  map-making that enables data reduction on single, modern, high-end
  desktop computers, with execution times that are typically shorter
  than the observing times.  SMURF is used in an automated setting,
  both at the telescope for real-time feedback to observers, as well
  as for the production of science products for the JCMT Science
  Archive at the Canadian Astronomy Data Centre. Three detailed case
  studies are used to: (i) explore convergence properties of the
  map-maker using simple prior constraints (Uranus -- a point source);
  (ii) achieve the white-noise limit for faint point-source studies
  (extragalactic blank-field survey of the Lockman Hole); and (iii)
  demonstrate that our strategy is capable of recovering angular
  scales comparable to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $10,240$ nominal detectors \citep[of which $\sim$70\%
work and are typically useful,][]{holland2012}, \scuba\ is presently
the largest of a new generation of multiplexed kilopixel
(sub)millimetre bolometer cameras, which also includes the cameras for
the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
experiments for studying anisotropies in the Cosmic Microwave
Background (CMB) using similar technology, the latter of which uses the
same time-domain multiplexed readout electronics as \scuba\
\citep{battistelli2008}.

Submillimetre imaging cameras generally maximise sensitivity using
bolometers, rather than coherent detectors, which are limited by white
photon and phonon noise from the instrument and ambient
backgrounds. The low-frequency noise, however, is typically dominated
by sources which produce slow variations in the background (e.g.,
thermal variations within the cryostat, and the atmosphere for
ground-based cameras), and drifts in the readout electronics. Such
noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
frequency at which it is comparable to the white noise level is called
the ``$1/f$ knee''. Since the low-frequency noise is largely
correlated between all, or subsets, of the bolometers in time, it can
be suppressed during map-making, since astronomical signals have the
distinct property that they are fixed in a sky reference frame
(assuming they are not time-varying). If successful, the noise in the
resulting map is said to be ``white noise limited'', meaning that it
is uncorrelated spatially, and has an amplitude that scales as the
$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
flux-density (the white noise level of a bolometer in 1\,s of
integration), and $t$ is the amount of integration time in a map
pixel.

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is scanned at a
range of position angles, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$), such that significant drift in the
bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
preference to ``chopping'' methods (where the secondary is moved
quickly to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensitive to modest angular scales. We note that under some
circumstances cross-linking may not be essential. For example, SPT,
which was designed to measure CMB anisotropies, scans entirely in
azimuth \citep{schaffer2011}. While this strategy limits its ability
to recover large angular scales transverse to the scan direction, the
anisotropic filtering in its maps can be accounted for during analysis
to achieve SPTs focussed objectives \citep[in contrast, the similar
ACT experiment uses cross-linking to improve its response to
large-angular scales,][]{das2011}. This approach is less practical as
a general solution for \scuba, however, which must serve a broader
range of scientific interests.

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same generation. A
more recent example is the analysis of SPT data
\citep{schaffer2011}. Generally speaking, such methods are fast,
although depending on the science goals and noise properties of the
data, they may not achieve the best noise performance on the angular
scales of interest. A related method that has been popular for
reducing data in fields of faint point sources with Bolocam
\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
principal component analysis (PCA) cleaning. A statistical
``black-box'' removes the most correlated components of the bolometer
signals, enabling the detection of point-sources close to the
theoretical white-noise limits of the detectors, with reasonable
computation times when small numbers of bolometers are involved
(hundreds rather than thousands of detectors). However, PCA cleaning
is not a good solution for producing maps of extended structures,
since such sources produce correlated signals amongst many detectors,
and are removed by this procedure \citep[an exception is the iterative
PCA approach of][]{aguirre2011}. Furthermore, in the case of \scuba,
performing PCA on even a single subarray (typically $\sim$900
detectors) can be prohibitively slow.

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the experimental CMB literature
include \citet{tegmark1997} and \citet{stompor2002}, and an
application to data from SCUBA is described in \citet{borys2004}.  The
downside to such methods is that they can be both computationally
expensive and have large memory requirements. While for some
experimental designs fast iterative methods for the inversion do exist
without requiring excessive amounts of memory, such as that of
\citet{wright1996} \citep[which was later implemented for SCUBA
by][]{johnstone2000}, for the more general map-making problem,
involving many detectors, matters are significantly complicated both
by the need to measure the cross spectra for all unique pairs of
bolometers, as well as performing the inversion itself. A
maximum-likelihood method was successfully used to make maps from ACT
data using approximately the same number of detectors as a single
\scuba\ subarray at a single wavelength. However, the calculation is
extremely resource intensive and requires a large cluster in order to
run; a single map can take 10\,CPU\,years \citep{fowler2010}. Perhaps
the most promising maximum-likelihood method that may one day be
applied to \scuba\ data is ``SANEPIC'', which produced maps from
Balloon-borne Large-Aperture Submillimeter Telescope data, involving
hundreds of bolometers, while correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the inversion step
that is so complicated in the maximum-likelihood methods. In the
iterative approach memory usage usually scales linearly with $N$, the
product of the number of bolometers, $N_\mathrm{b}$, with the number
of samples in time, $N_\mathrm{t}$. The computation time is the
product of the number of iterations, $N_\mathrm{iter}$, with the time
per iteration, $t_\mathrm{iter}$. For the algorithm described in this
paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
(depending on the \snr\ of large-scale structures, and the size of the
map), and $t_\mathrm{iter}$ scales as $N_\mathrm{t} \log
N_\mathrm{t}$, since Fast Fourier Transforms (FFTs) of the time-series
are typically performed. By comparison, the fastest maximum-likelihood
methods that account for inter-bolometer correlations have an
execution time with an $N_\mathrm{b}^2$ dependence, although it is
possible to limit memory usage to be linear in $N_\mathrm{t}$
\citep[e.g.,][]{patanchon2008}.

Iterative approaches to map-making have a long history in the
submillimetre, such as the pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The closest modern
relatives are the Comprehensive Reduction Utility
\citep[CRUSH,][]{kovacs2008} for the Submillimeter High-Angular
Resolution Camera 2 (SHARC-2), the pipeline developed for the Bolocam
Galactic Plane Survey \citep{aguirre2011}, and the Bolometer array
data Analysis software (BoA) for the Large APEX BOlometer CAmera
\citep[LABOCA,][]{schuller2012}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales commensurate with the observing
times (or significantly faster), on single high-end desktop computers
using the . For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline \citep[based on the
Observatory Reduction and Acquisition Control - Data Reduction system,
or ORAC-DR,][]{gibb2005} offering feedback to observers at the
telescope. The pipeline is also used to generate products for the JCMT
Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
Astronomy Data Centre (CADC). Details of the pipeline design are given
in \citet{gibb2005} and \citet{2008AN....329..295C}.

@@

In this section we summarise how \scuba\ works,
(Section~\ref{sec:bolos}), give examples of typical bolometers signals
(Sections~\ref{sec:bolosignal}), discuss the impacts of magnetic field
pickup (Section~\ref{sec:magpickup}) and sky noise
(Section~\ref{sec:skynoise}), and finally use of principal component
analysis to explore the correlated noise properties of \scuba\ data
(Section~\ref{sec:pca}).

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers (with a response
time constant of $\tau \sim 1$\,ms) coupled to superconducting
transition-edge sensors (TESs) for thermometry. Temperature variations
in the TESs produce changing currents, and therefore varying magnetic
fields, which are detected and amplified using chains of
superconducting quantum interference devices (SQUIDs), before the
larger output currents are digitised. Each detector has its own SQUID
for the first-stage of the amplification, but the remaining stages
occur within a common chain of SQUIDs for each of the 32 columns. All
40 rows are read out in sequence at a row-visit frequency of about
12\,kHz. Such a high sample rate is unnecessary to produce maps, so
the data are re-sampled to approximately 200\,Hz before writing to
disk. This rate provides a sample every 3.0\,arcsec, or approximately
one third of the 450\,\micron\ diffraction-limited full-width at
half-maximum (FWHM) -- a typical rule-of-thumb for adequately sampling
a Gaussian point spread function (PSF) -- at the maximum scanning
speed of 600\,arcsec\,s$^{-1}$. There is an additional 41st row of
SQUID readouts that are not connected to TESs. These ``dark SQUIDs''
track non-thermal noise sources that are common to each column's
amplifier chain. The relationship between the output digitised
current, $I$, and the input power, $P$, or $dI/dP$ is established
using flatfield observations immediately prior to science
observations, in which the output signal is measured throughout a ramp
of the pixel heaters (which provide a known input power); see
Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012} for more details. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the 183\,GHz (1.6\,mm)
JCMT Water Vapour Monitor to track line-of-sight opacity variations,
see Section~3 in][]{dempsey2012}, and the application of a flux
conversion factor (FCF) which is established from regular measurements
of calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
thermal time constant, as well as a negative shift of about $-$2.5\,ms
caused by a half-sample offset due to the synchronization between the
pointing and bolometer data, yields a net lag of approximately
4.5\,ms. There are also line features in the spectra that are thought
to be produced primarily at frequencies far above the final 200\,Hz
sample rate, which are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area at different
scan angles); and (ii) minimal accelerations to reduce turn-around
overheads (which could be quite large given the 600\,arcsec\,s$^{-1}$
maximum scan speed). For areas larger than the array footprint a
rectangular ``PONG'' pattern is used, in which the boresight travels
in approximately straight lines and ``bounces'' off the edges at
45\,degree angles until the area is uniformly filled in. It is also
usually combined with a rotation through a number of fixed position
angles to create a ``rotating PONG'' with even better
cross-linking. For smaller areas (of order the array footprint, or
point sources), in which the PONG turn-around overheads would be
large, a ``constant velocity daisy'' (CV Daisy) is used.  Here the
telescope moves in a circle, whose centre also slowly traces out a
small circle. For a more complete description of the \scuba\ observing
modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

\begin{figure*}
\centering

\begin{minipage}[h]{0.495\linewidth}
\textbf{(a) S2SRO} \\

\includegraphics[width=\linewidth]{bolos_point_mix_s2sro}
\end{minipage}
\begin{minipage}[h]{0.495\linewidth}
\textbf{(b) fully-commissioned} \\

\includegraphics[width=\linewidth]{bolos_point_mix}
\end{minipage}

\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature,
  the line-of-sight 450\,\micron\ opacity derived from the JCMT water
  vapour monitor, and azimuth/elevation pointing offsets, before and
  after upgrading the instrument. The grey signals over-plotted in the
  top two panels show the residual time-series after removing the
  common-mode signals.  (a) Data taken during the S2SRO period,
  observation 29 on 20100313. There is a strong correlation between
  the bolometers and the roughly $\sim$25\,s oscillation in the
  fridge, but only a minor correlation with the opacity and telescope
  motion. The nearly flat common-mode subtracted signals show: (i)
  that most of the low-frequency signal is common to all of the
  bolometers; and (ii) the non-correlated, and predominantly white
  noise at 450\,\micron\ is significantly larger than at
  850\,\micron. (b) Data taken with the fully-commissioned instrument,
  observation 38 on 20111112. Unlike the S2SRO data, there is no
  strong signal produced by variations in the fridge temperature. The
  450\,\micron\ data have minimal correlated low-frequency noise (as
  evidenced by the lack of a difference between the raw and
  common-mode subtracted signals), although the 850\,\micron\ data
  show a common signal correlated with the opacity and telescope
  motion (note that this scan has a significantly larger amplitude
  than the S2SRO data set), mostly likely caused by a combination of
  changing airmass (anti-correlated with opacity), and magnetic field
  pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

\begin{figure*}
\centering

\begin{minipage}[h]{0.495\linewidth}
\textbf{(a) S2SRO} \\

\includegraphics[width=\linewidth]{pspec_s2sro}
\end{minipage}
\begin{minipage}[h]{0.495\linewidth}
\textbf{(b) fully-commissioned} \\

\includegraphics[width=\linewidth]{pspec}
\end{minipage}

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centred over 450 and 850\,\micron. The camera is
  operated by scanning across the sky and records data at a rate of
  200\,Hz. This rate is much greater than that of previously existing
  submillimetre cameras, and represents a significant analysis
  challenge. We describe the production of \scuba\ maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products for the JCMT Science Archive at the Canadian
  Astronomy Data Centre. Three detailed case studies are used to: (i)
  explore convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $\sim10,000$ total detectors, \scuba\ is presently
the largest submillimetre camera in the world, and the fastest
wide-area ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers, rather than
coherent detectors, to maximise sensitivity. The sensitivity of
bolometers is limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is typically dominated by sources which produce slow variations in the
background (e.g., thermal variations within the cryostat, and the
atmosphere for ground-based cameras), and drifts in the readout
electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
($\alpha>0$), and the frequency at which it is comparable to the white
noise level is called the ``$1/f$ knee''. Since the low-frequency
noise is largely correlated between all, or subsets, of the bolometers
in time, it can be suppressed during map-making, since astronomical
signals have the distinct property that they are fixed in a sky
reference frame (assuming they are not time-varying). If successful,
the noise in the resulting map is said to be ``white noise limited'',
meaning that it is uncorrelated spatially, and has an amplitude that
scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
noise-equivalent flux-density (the white noise level of a bolometer in
1\,s of integration), and $t$ is the amount of integration time in a
map pixel.

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$), such that significant drift in the
bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
preference to ``chopping'' methods (where the secondary is moved
quickly to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensitive to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
depending on the science goals and noise properties of the data, they
may not achieve the best noise performance on the angular scales of
interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with Bolocam
\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
``principal component analysis'' (PCA) cleaning. A statistical
``black-box'' removes the most correlated components of the bolometer
signals, enabling the detection of point-sources very close to the
theoretical white-noise limits of the detectors, with reasonable
computation times when small numbers of bolometers are involved
(hundreds rather than thousands of detectors). However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since such sources produce correlated signals amongst many
detectors, and are removed by this procedure. Furthermore, in the case
of \scuba\, performing PCA on even a single subarray (1280 detectors,
nominally) can be prohibitively slow.

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the experimental cosmic microwave
background (CMB) literature include \citet{tegmark1997} and
\citet{stompor2002}, and an application to data from SCUBA is
described in \citet{borys2004}.  The downside to such methods is that
they can be both computationally expensive and have large memory
requirements. While for some experimental designs fast iterative
methods for the inversion do exist without requiring excessive amounts
of memory \citep[e.g.,][]{wright1996}, for the more general map-making
problem, involving many detectors, things are significantly
complicated both by the need to measure the cross power spectra for
all unique pairs of bolometers, as well as performing the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, while correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the measurement of
noise properties and the inversion step that is so complicated in the
maximum-likelihood methods. In principle, it should also be possible
to recover a wider dynamic range in angular scales than possible for
the simplest direct methods, with an increase in calculation time that
is merely linear in the number of iterations needed, while using a
comparable amount of memory. The iterative method of
\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
observations by \citet{johnstone2000}, although in this case the
chopping removes most of the low-frequency noise, and greatly
simplifies the problem (at the expensive of large-scale structure in
the data). More similar to the algorithm described here was the
iterative pipeline recipe for fitting and removing baseline drifts in
SCUBA scan-map data as the astronomical image estimate improved
\citep{1999ASPC..172..171J}. The closest modern relatives of SMURF are
the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
Bolometer array data Analysis software (BoA) for the Large APEX
BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
developed for the Bolocam Galactic Plane Survey \citep[][which uses
iterative PCA cleaning to avoid the suppression of large-scale
structure]{aguirre2011}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data faster than they are taken, on single high-end
desktop computers, and it has successfully been used as part of a
real-time pipeline offering feedback to observers at the
telescope. The pipeline is also used to generate near-science grade
products for the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomy Data Centre (CADC). Details of the
pipeline design are given in \citet{gibb2005} and
\citet{2008AN....329..295C}.

@@

In this section we summarise how \scuba\ works,
(Section~\ref{sec:bolos}), give examples of typical bolometers signals
(Section~\ref{sec:bolosignal}), and the impact of magnetic field
pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
of principal component analysis to explore the correlated noise
properties of \scuba\ data (Section~\ref{sec:pca}).

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry.
Temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields, which are detected and amplified
using chains of superconducting quantum interference devices (SQUIDs),
before the larger output currents are digitised. Each detector has its
own SQUID for the first-stage of the amplification, but the remaining
stages occur within a common chain of SQUIDs for each of the 32
columns. All 40 rows are read out in sequence at a row-visit frequency
of about 12\,kHz. Such a high sample rate is unnecessary to produce
maps, so the data are re-sampled to approximately 200\,Hz before
writing to disk. This rate provides a sample every 3.0\,arcsec, or
approximately one third of the 450\,\micron\ diffraction-limited
full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
adequately sampling a Gaussian point spread function (PSF) -- at the maximum scanning speed. There is
an additional 41st row of SQUID readouts that are not connected to
TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
common to each column's amplifier chain. The relationship between the
output digitised current, $I$, and the input power, $P$, or $dI/dP$ is
established using flatfield observations immediately prior to science
observations, in which the output signal is measured throughout a ramp
of the pixel heaters (which provide a known input power); see
Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012} for more details. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the JCMT Water Vapour
Monitor to track line-of-sight opacity variations, see Section~3
in][]{dempsey2012}, and the application of a flux conversion factor
(FCF) which is established from regular measurements of calibrators
such as Uranus \citep[Section~5 in][]{dempsey2012}.

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra include a roll-off in their spectra at frequencies approaching
Nyquist, which is due to the anti-aliasing filter that forms part of
the 200\,Hz re-sampling stage. There are also line features which are
thought to be produced at frequencies far above the final 200\,Hz
sample rate, and are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area on several
different time scales, and at different scan angles); and (ii) minimal
accelerations to reduce turn-around overheads (which could be quite
large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
larger than the array footprint a rectangular ``PONG'' pattern is
used, in which the boresight travels in approximately straight lines
and ``bounces'' off the edges at 45\,degree angles until the area is
uniformly filled in. It is also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking. For smaller areas (of order the
array footprint, or point sources), in which the PONG turn-around
overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
used.  Here the telescope moves in a circle, whose centre also slowly
traces out a small circle. For a more complete description of the
\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
\includegraphics[width=0.49\linewidth]{bolos_point_mix}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals over-plotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{pspec_s2sro}
\includegraphics[width=0.49\linewidth]{pspec}
\caption{Bolometer power spectral densities for the same two data sets
  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
  noise slightly and clarify some features. Four of the most sensitive
  bolometers have been selected at each wavelength, and the
  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
  as coloured dotted lines (the blue signals are for the same time
  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
  lines are the PSDs of the common-mode signals at each wavelength,
  and the solid coloured lines show the PSDs of the bolometers once
  the common-mode is removed.  Finally, the dashed black lines show
  the spectral shape produced by a point source given the scan speeds
  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
  reference, the top horizontal axes shows the conversion from
  frequency to angular scale. Horizontal dotted lines at $2 \times
  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
  respectively, are provided as a visual reference for the white-noise
  levels. In addition to $1/f$ and white noise components, and line
  features, all of the PSDs exhibit the gradual roll-off of the
  anti-aliasing filter just below the Nyquist frequency.  Left: for
  the S2SRO data there are broad line features in the PSDs at both
  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
  450 and 850\,\micron. Common-mode subtraction removes most of the
  correlated fridge oscillation signal, lowering the $1/f$ knees to
  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
  respectively. Right: data from the fully-commissioned instrument
  tend to have lower $1/f$ knees, and fewer line features. The
  remaining low-frequency noise, now that the fridge oscillations have
  been removed, is considerably more independent from bolometer to
  bolometer than in the S2SRO data (note the larger spread in the
  dotted lines at low frequencies compared to the left-hand plot,
  particularly at 450\,\micron). Finally, this example shows that the
  white noise performance is in fact similar for the two subarrays
  (s4a and s8b) that were used both before and after the upgrades.}
\label{fig:pspec}
\end{figure*}
